---
title: "cdsrgsea"
output:
  github_document:
  html_notebook:
    theme: united
  html_document:
    theme: united
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 7, fig.height = 7, cache = T)

library(tidyverse)
library(useful)
library(taigr)
library(cdsrgsea)
library(cdsrmodels)
```

cdsrgsea contains lightweight wrappers around useful gene set enrichment functions

## Install

```{r, eval = FALSE}
library(devtools)
devtools::install_github("broadinstitute/cdsr_gsea")
```

The package can then be loaded by calling

```{r, eval=FALSE}
library(cdsrgsea)
```

```{r include = F}
gene_sets <- cdsrgsea::load_gene_sets()
diff_expr <- read_csv("./nutlin.csv")
```

## load_gene_sets

Loads a list of gene sets in term2gene format from taiga. These gene sets are drawn from the [Enrichr](https://amp.pharm.mssm.edu/Enrichr/#stats) and [MSigDB](https://www.gsea-msigdb.org/gsea/msigdb/index.jsp) libraries.

```{r eval = F}
gene_sets <- cdsrgsea::load_gene_sets()
```

```{r}
gene_sets %>% names() %>% head()
```

```{r}
gene_sets$Hallmark %>% head()
```

## run_hyper

Runs overrepresentation analysis based on the hypergeometric distribution. The function excepts either a small list of significant genes or a table with gene level stats as input. If a table is provided the `gene_var`, `rank_var`, and `n_genes` parameters are used to define a small list of significant genes.

1. Small list of significant genes

```{r}
genes <- diff_expr %>% arrange(logFC) %>% tail(100) %>% .[["Gene"]]
genes %>% head()
```

```{r}
hyper_res <- cdsrgsea::run_hyper(genes,gene_sets$Hallmark)
hyper_res
```

1. A table with gene level stats

```{r}
ora_res <- cdsrgsea::run_hyper(diff_expr,gene_sets$Hallmark,gene_var = "Gene",rank_var = "logFC",dir = "pos")
ora_res
```

## run_gsea

Runs gene set enrichment analysis using the `fgseaMultilevel` method from the `fgsea` package. The function excepts a table with gene level stats. The `dir` parameter indicates whether to return positive gene sets, negative gene sets, or both.

```{r}
gsea_res <- cdsrgsea::run_gsea(diff_expr,gene_sets$Hallmark,gene_var = "Gene",rank_var = "logFC",dir = "neg")
gsea_res
```






